---
title: "Getting started"
output:
  rmarkdown::html_vignette:
    md_extensions: -smart
    highlight_downlit: true
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  message = FALSE,
  warning = FALSE,
  dev = "svglite",
  fig.width = 7,
  fig.height = 9,
  fig.align = "center"
)
```

```{css, echo=FALSE}
.sourceCode {
  overflow: auto !important;
}
```

```{r setup}
library(hBayesDMregression)

# Tidyverse packages
library(tibble)
library(dplyr)
library(ggplot2)
library(tidyr)
library(purrr)
library(magrittr)

# For MCMC diagnostics
library(rstan)
```

# An example model

```{r}
data("igt_example")

head(igt_example)
```

Note that the included example data set has a `trial` column that records the trial number for each
subject. This column should be excluded when building the model.

```{r, eval=FALSE}
## NOT RUN
orl_model <- igt_orl_regression(
  igt_example,
  exclude_cols = "trial",
  regression_pars = c("K", "betaF"),
  nchain = 3,
  niter = 3000,
  nwarmup = 1000,
  ncore = parallel::detectCores() / 2
)
```

```{r}
orl_model <- readRDS("orl_model.rds")
```

# Distributions of posterior samples

The posterior samples are contained in the `parVals` index of the model output. We will extract and
assign it to a variable in our workspace since we will be calling it often.

```{r}
posterior_samples <- orl_model %>%
  pluck("parVals")
```

Note: the above is equivalent to calling either `orl_model$parVals` or `orl_model[["parVals"]]`.

```{r}
str(posterior_samples)
```

## Hyperparameters

### Mu

```{r}
mu <- posterior_samples %>% # This is a list
  keep_at(~ grepl("^mu_", .x)) %>% # Keep list elements whose names begin with `mu_`
  do.call(cbind, .) %>% # Column-bind the list elements
  as_tibble() %>% # Matrix -> Tibble
  pivot_longer(everything(), names_to="parameter") # Reshape to long format

mu
```

```{r}
ggplot(mu) +
  geom_histogram(aes(x=value, y=after_stat(density)), colour="#56B4E9", fill="#0072B2", alpha=0.7) +
  facet_wrap(~parameter, scales="free", ncol=2) +
  theme_bw() +
  labs(x="Value", y="Density")
```

### Sigma

```{r}
sigma <- posterior_samples %>% # This is a list
  pluck("sigma") %>% # This is a matrix
  set_colnames(paste0("sigma_", orl_model$modelPars)) %>% # Set column names
  as_tibble() %>% # Matrix -> Tibble
  pivot_longer(everything(), names_to="parameter") # Reshape to long format

sigma
```

```{r}
ggplot(sigma) +
  geom_histogram(aes(x=value, y=after_stat(density)), colour="#56B4E9", fill="#0072B2", alpha=0.7) +
  facet_wrap(~parameter, scales="free", ncol=2) +
  theme_bw() +
  labs(x="Value", y="Density")
```

## Model parameters

```{r}
model_parameters <- posterior_samples %>% # This is a list
  keep_at(orl_model$modelPars) %>% # This is a list of matrices
  map(~ set_colnames(.x, paste0(orl_model$subjID))) %>% # Set column names for each matrix in the list
  map(as_tibble) %>% # Matrix -> Tibble for each matrix in the list
  map(~ pivot_longer(.x, everything(), names_to="subjID")) # Reshape each matrix within list to long format

model_parameters
```

```{r}
ggplot(model_parameters$Arew) +
  geom_histogram(aes(x=value, y=after_stat(density)), colour="#56B4E9", fill="#0072B2", alpha=0.7) +
  facet_wrap(~subjID, scales="free", labeller="label_both") +
  theme_bw() +
  labs(x="Value", y="Density", caption="Model parameter Arew by subject")
```

## Regression parameters

### Beta

```{r}
beta <- posterior_samples %>%
  pluck("beta") %>% # This is a matrix and already has column names
  as_tibble() # Matrix -> Tibble

K_betas <- beta %>%
  select(starts_with("K")) %>% # Select betas corresponding to model parameter K
  pivot_longer( # Reshape to long format
    everything(),
    names_to = "parameter",
    names_transform = ~ sub("K_", replacement="", .x)
  )

K_betas
```

```{r}
ggplot(K_betas) +
  geom_histogram(aes(x=value, y=after_stat(density)), colour="#56B4E9", fill="#0072B2", alpha=0.7) +
  facet_wrap(~parameter, scales="free") +
  theme_bw() +
  labs(x="Value", y="Density", caption="Betas for model parameter K")
```

### Sigma beta

```{r}
sigma_beta <- posterior_samples %>%
  pluck("sigma_beta") %>% # This is a matrix and already has column names
  as_tibble() # Matrix -> Tibble

K_sigma_betas <- sigma_beta %>%
  select(starts_with("sigma_K")) %>% # Select sigma betas corresponding to model parameter K
  pivot_longer( # Reshape to long format
    everything(),
    names_to = "parameter",
    names_transform = ~ sub("sigma_K_", replacement="", .x)
  )

K_sigma_betas
```

```{r}
ggplot(K_sigma_betas) +
  geom_histogram(aes(x=value, y=after_stat(density)), colour="#56B4E9", fill="#0072B2", alpha=0.7) +
  facet_wrap(~parameter, scales="free") +
  theme_bw() +
  labs(x="Value", y="Density", caption="Sigma betas for model parameter K")
```

# Traceplots

Since the original Stan fit is included in the `hBayesDMregression` model output, we can pass
it directly to `rstan::traceplot()` to create traceplots. The parameters whose traceplots can be
plotted can be found by getting the names of the Stan fit:

```{r}
names(orl_model$fit)
```

Suppose we want the traceplots for all parameters related to covariate `x4`:

```{r}
pars_x4 <- grep("x4$", names(orl_model$fit), value=TRUE)

pars_x4
```

```{r}
traceplot(orl_model$fit, pars=pars_x4)
```

Note that `rstan::traceplot()` uses `{ggplot2}` for plotting. As such, additional customizations can
be made via the `{ggplot2}` package.

# Interval plots

Intervals plots are constructed similarly.

```{r}
pars_age <- grep("age$", names(orl_model$fit), value=TRUE)

pars_age
```

```{r}
stan_plot(orl_model$fit, pars=pars_age)
```

Note that `rstan::stan_plot()` uses `{ggplot2}` for plotting. As such, additional customizations can
be made via the `{ggplot2}` package.
